Avoidance in permutations

### 2014-11-07 project2/paper/paper.tex
added flipping and reversing to section5

\section{Terms in the sequence $s_n(321)$}
\label{321}
Unfortunately, there is no clear way to reformulate Lemma \ref{lemma1} for the case of avoiding the permutation 321 or 123. As a consequence, the proof in the previous section simply does not hold. Fortunately, however, we still have a closed form expression for $s_n(321)$ (recall that $s_n(321)=s_n(123)$ by our flipping principle). Surprisingly, in fact, the sequence remains the same: $s_n(321)=C_n$, the $n^{th}$ Catalan number. In order to prove this result, we need the machinery of Young tableaux. What follows is a brief exposition of the key results, adapted from Stanley's \emph{Algebraic Combinatorics}.

@@

In this section we present our findings and conjectures about sequences $s_n(\pi)$ for $\pi\in S_4$. We computed values up to $s_{9}$. It appears that there are three sequences that appear, and we have sorted the permutations according to which sequence they produce and paired them with their flips.

@@

There is no guarantee that these sequences will not diverge and produce more than three distinct sequences, however for now we hypothesize that $A$, $B$, and $C$ are the only sequences that arise for $s_n(\pi)$ with $\pi\in S_4$. Assuming this hypothesis, if we write each length four permutation in cycle notation, a pattern begins to emerge.

\begin{center}
\begin{tabular}{|c|c|c|}
B &A&C\\
\hline
(1),(14)(23)&(243),(142)&(23),(14)\\
(34),(1423)&(234),(143)&\\
(24),(1432)&(124),(132)&\\
(12),(1324)&(123),(134)&\\
(12)(34),(13)(24)&(1243),(1342)&\\
(1234),(13)&&\\
\end{tabular}
\end{center}

So we conjecture that sequence $A$ is characterized by elements of $S_4$ that are three cycles or four cycles when written in cycle notation. Sequence $B$ seems to be characterized by either a pair of disjoint two cycles paired with a flip that is a 2-cycle, or a single 4-cycle paired with a flip that is a 2-cycle. Sequence $C$ appears to be characterized by a 2-cycle paired with a 2-cycle as a flip.

>>>>

\section{Terms in the sequence $s_n(321)$}

\label{321}
Unfortunately, there is no clear way to reformulate Lemma \ref{lemma1} for the case of avoiding the permutation 321 or 123. As a consequence, the proof in the previous section simply does not hold. Fortunately, however, we still have a closed form expression for $s_n(321)$ (recall that $s_n(321)=s_n(123)$ by our flipping principle). Surprisingly, in fact, the sequence remains the same: $s_n(321)=C_n$, the $n^{th}$ Catalan number. In order to prove this result, we need the machinery of Young tableaux. What follows is a brief exposition of the key results, adapted from Stanley's \emph{Algebraic Combinatorics}.

@@

In this section we present our findings and conjectures about sequences $s_n(\pi)$ for $\pi\in S_4$. Using the flipping and reversing operators we can group elements of $S_4$ by their orbits. The following groups appear: 
$$
\{1234,4321\},
\{1243,4312,2134,3412\},
\{1432,4123,2341,3214\},
\{2143,3412\},
$$
$$
\{4132,1423,2314,3241\},
\{4213,1342,3124,2431\},
\{2413,3142\}
$$
$$
\{4231,1324\}
$$
So all permutations that belong to a single group in the list above must produce the same sequence. 
We computed values up to $s_{9}$. It appears that there are three sequences that appear, and we have sorted the permutations according to which sequence they produce and paired them with their flips.

@@

Comparing this table to the orbit decomposition above we notice that all of the groups in the first line appear to produce sequence $B$, all of the groups in the second line appear to produce sequence $A$, and the single orbit in line three produces sequence $C$. There is no guarantee that these sequences will not diverge and produce more than three distinct sequences, however for now we hypothesize that $A$, $B$, and $C$ are the only sequences that arise for $s_n(\pi)$ with $\pi\in S_4$. Assuming this hypothesis, if we write each length four permutation in cycle notation, a pattern begins to emerge.

\begin{center}
\begin{tabular}{|c|c|c|}
B &A&C\\
\hline
(1)(2)(3)(4),(14)(23)&(243),(142)&(23),(14)\\
(34),(1423)&(234),(143)&\\
(24),(1432)&(124),(132)&\\
(12),(1324)&(123),(134)&\\
(12)(34),(13)(24)&(1243),(1342)&\\
(1234),(13)&&\\
\end{tabular}
\end{center}

So we conjecture that sequence $A$ is characterized by elements of $S_4$ that are three cycles or four cycles when written in cycle notation. Sequence $B$ seems to be characterized by either a pair of disjoint two cycles paired with a flip that is a pair of disjoint 2-cycles, or a single 4-cycle paired with a flip that is a 2-cycle. Sequence $C$ appears to be characterized by a 2-cycle paired with a 2-cycle as a flip.
